Spin canonical divisors of log stacky curves

### 2015-06-30 stacky_write_up/spin_cring.tex
david's comments

\todo{combine below two paragraphs with the corresponding
near-duplicate part from the background}
If $X$ is a curve, \todo{why do you start with a curve and then
move to a stacky curve? shouldn't it be the opposite order?} then
we say $\sx$ is a {\bf stacky curve} with {\bf coarse space} $X$
and {\bf stacky points} (also called {\bf{fractional points}}) $P_1,
\ldots, P_r$ with {\bf stabilizer orders} $e_1,\ldots, e_r \in \BZ_{
\geq 2}$ if 
\[
	\di \sx = \di X \oplus \left(\bigoplus_{i = 1}^r \left\langle 
	\frac{1}{e_i}P_i \right\rangle \right) \subset \BQ \otimes \di X.
\]
\todo{Peter: Would it be better to say $Div(\sx)=(Div(X))[\frac{1}{e_i}P_i]_{1\le i\le r}$ or something like that? Aaron: I agree what is written isn't actually correct, maybe we should say the free direct sum over points not in the $P_i$, together with the fractional points. It's not a localization because we don't want to invert $P_i$.}
As is done in Voight and Zureick-Brown \cite{vzb:stacky} stacky curves may be formalized as stacks, but for the purposes of this paper, it suffices to view $\sx$ as a scheme whose associated divisor group allows a finite number of stacky points. \todo{I don't know if this is a good statement to make.  We aren't really viewing $\sx$ as a scheme.  Rather we're are determining information about its canonical rings through knowledge of scheme data.}  A {\bf log spin curve} is a triple $(\sx, \Delta, \halfcan)$ where $\Delta \in \di X$ is an effective divisor and $2\halfcan = K_X + \Delta + \sum_{i=1}^{r}\subhalf{e_i}P_i.$
The central object of study in this paper is the {\bf log spin
canonical ring} of $(\sx, \Delta, \halfcan)$ is
\begin{align*}
	R(\sx, \Delta, \halfcan) = \bigoplus_{d \geq 0} H^0(\sx,\lfloor d \halfcan \rfloor)
\end{align*}

@@

The goal of this paper is to provide a stricter bound on the
generator and relation degrees of log spin canonical rings
for all genera. Note that O'Dorney's bounds are asymptotically
tight for arbitrary $\BQ$-divisors on $\BP^1$, but restricting
to the log spin curve case allows for lower bounds.
The main theorem of this paper is the following improved bound on
the degree of generators and relations of log spin canonical rings:

\todo{State for even weight modular forms attributed to VZB.}
\begin{thm}
\label{thm:main}
Let $(\sx, \Delta, \halfcan)$ be a log spin curve over a perfect field $k$, \todo{Ask David: do we need this perfectness assumption everywhere?} so
that $\sx$ has signature $\sigma = (0; e_1, \ldots, e_r; \delta)$. Then the
canonical ring

@@

\ssec{Stacky Curves and Log Spin Canonical Rings}
\label{ssec:stacky-background}
\todo{combine the below chunk with the near-duplicate section
in the introduction}
A \textbf{stacky curve} $\sx$ over a field $k$ is a smooth proper
geometrically-connected Deligne-Mumford stack of dimension $1$ over
$k$ with a dense open subscheme. A stacky curve is \textbf{tame} if
its stabilizers are coprime to char $k$. Throughout this paper,
stacky curves will be assumed to be tame and $k$ will always be
algebraically closed. The language of stacks will be omitted in
this paper, as it is sufficient to view stacky curves and related
objects $\sx$ as a scheme whose associated divisor group allows for
a finite number of \textbf{stacky points} which may be viewed
heuristically as ``fractional points''. See Voight and
Zureick-Brown for more thorough treatment of these definitions
\cite[Chapter 5]{vzb:stacky}. \todo{again, not quite viewing as a scheme}

@@

\noindent
When the log spin curve is fixed, we usually use $R$ or $R_L$ to
represent $R(\sx, \Delta, \halfcan)$. We also often use $H^0(\sx, D)
:= H^0(\sx, \sco(D))$ and $h^0(\sx, \sco(D)) := \dim H^0(\sx, \sco(D))$.
If $R$ is a graded ring, then we let $(R)_d$ refer to the $d^{th}$
degree component of $R$.

@@

@@

\begin{lem}
\label{lem:sat-2}
Let $(\sx, \Delta, \halfcan_\sx)$ and $(\sx', \Delta, \halfcan_{\sx'})$ \todo{The notation i was using for these throughout the paper was $\halfcan$ and $\halfcan'$ without the $\sx$ subscripts. We should make one of them consistent with the other.} be log spin curves with the same coarse space $X = X'$ having signatures $(g; e_1, \ldots, e_r; \delta)$ and $(g, e_1, \ldots, e_{r- 1}, \delta)$, where $e_r=3$.  Suppose $g > 0,$ and, if $g = 1,$ then $\deg 3\halfcan' \geq 2.$ Then, $\sat(Eff(L'))\le 2$. \todo{Why is it automatic that the saturation is 2? Maybe you can explain this in the proof in a sentence.}
Furthermore, let $R_{L'} = k[x_2, x_3
, x_5, \ldots, x_m]/I'$ and let $\halfcan = \halfcan' + \frac{1}{
3}P$, where $P\in X$ is a base point of $\halfcan'$ (which includes the case when $H^0(X, \lfloor \halfcan'\rfloor) = 0$). \todo{Ask David, is it useful to include this parenthetical?}
Suppose for $i \in \{2, 3\}$, $\deg x_i = i$ and $\ord_P^{\halfcan'}(x_i)= 0$ (which is satisfied generically in degree 2 and 3, since base-point freeness follows by Riemann-Roch \todo{Probably include this in a remark below the proof}). Choose an ordering on $k[x_2, \ldots, x_m]$ that satisfies
\begin{align*}
	\ord_{x_2}(f) < \ord_{x_2}(h) \implies f \prec h.
\end{align*}

@@

Therefore, general elements $y_3$ and $y_4$ satisfy $-\ord_P^{\halfcan'}(y_i) = 1$ by Riemann-Roch, proving part (a). \todo{Ask David, Understand why/whether we need to cite Lemma 5.4.7.}

@@

%\begin{rem}
%Note that $S_{\sigma,J}(i,d-1)$ as defined in condition (Ad-iii) of Definition ~\ref{defn:admissible} can equivalently be defined as $S_{\sigma,J}(i,d-1)= \{j \in J : j \neq i \text{ and }(e_j' + 2d-2)|e_i' + 2d\}.$ Written this way, $\# S_{\sigma,J}(i,d-1)$ is more apparently similar to $\mu(i,d)$ as defined in \cite[Defintion 8.5.1]{vzb:stacky}.
%\end{rem}
%\todo{Peter: I'm not sure how useful this remark is.  After seeing the definition of $S_{\sigma, J}(I,d)$ a reader will probably be a little mystified why it is a useful condition.  The inclusion of the relationship to something defined in VZB that also seems ad hoc and that they probably will not have seen may not be the most helpful use of this space.}

@@

\begin{lem}
\label{lem:genus-2-generation-5}
Let $X$ be a smooth genus 2 curve with canonical divisor $K$ and let $L \in \di X$ be so that $2L \sim K$. 
Then, $R_L$ is generated in degree at most $5$.
\end{lem}
\begin{proof}
By Lemma ~\ref{lem:semicanonical_generation} it suffices to show $R_L$ has no minimal generators in degree 6.
Let $P_1, \ldots, P_6$ be the six hyperelliptic fixed points of $X$. 
Recall that all linear equivalence classes of spin divisors on $X$ have a representative of the form
$$L = P_1 + \sum_{i =2}^{6} \epsilon_i (P_1 - P_i),$$ 
with $\epsilon_i \in \{0,1\},$ and not all $\epsilon_i = 1$. \todo{cite this somewhere, maybe ask david for a place to cite? We can cite http://www.math.harvard.edu/~ctm/home/text/class/harvard/213b/14/html/home/course/course.pdf, for example, from complex analysis course notes!}
Note that $H^0(X,K)$ is spanned by rational functions $u, x$ with $\di u|_{2L} = -2L, \di x|_{2L} = -2L + 2P_1,$ where $u,x$ are though of as rational functions. Therefore, the image of the multiplication map 
\begin{align*}
	Sym^3 H^0(X,2L) \rightarrow H^0(X,6L)
\end{align*}
is 4 dimensional by Riemann--Roch, with image spanned by $u^3,u^2x,ux^2, x^3$, 
which have 
\begin{align*}
	\di u^3|_{2L} &= -2L \\
	\di u^2x|_{2L} &= -2L + 2P_1 \\
	\di ux^2|_{2L} &= -2L+4P_1, \\
	\di x^3|_{2L} &= -2L + 6P_1.
\end{align*}

@@

\begin{align*}
	\dim_k (k[x_{3, 1}, x_{5, 1}, x_{7, 1}, x_{7, 2}] / I)_d &=
	\dim_k (k[x_{3, 1}, x_{5, 1}, x_{7, 1}, x_{7, 2}]	/ in_\prec(I))_d \\
	&= \dim_k (k[x_{3, 1}, x_{5, 1}, x_{7, 1}, x_{7, 2}] / \langle
	x_{3, 1} x_{7, 2}, x_{7, 2}^2 \rangle
\end{align*}

>>>>

If $X$ is a curve, \todo{why do you start with a curve and then
move to a stacky curve? shouldn't it be the opposite order?} then
we say $\sx$ is a {\bf stacky curve} with {\bf coarse space} $X$
and {\bf stacky points} (also called {\bf{fractional points}}) $P_1,
\ldots, P_r$ with {\bf stabilizer orders} $e_1,\ldots, e_r \in \BZ_{
\geq 2}$ if 
\[
	\di \sx = \di X \oplus \left(\bigoplus_{i = 1}^r \left\langle 
	\frac{1}{e_i}P_i \right\rangle \right) \subset \BQ \otimes \di X.
\]
\todo{Peter: Would it be better to say $Div(\sx)=(Div(X))[\frac{1}{e_i}P_i]_{1\le i\le r}$ or something like that? Aaron: I agree what is written isn't actually correct, maybe we should say the free direct sum over points not in the $P_i$, together with the fractional points. It's not a localization because we don't want to invert $P_i$.}
As is done in Voight and Zureick-Brown \cite{vzb:stacky} stacky curves may be formalized as stacks, but for the purposes of this paper, it suffices to view $\sx$ as a scheme whose associated divisor group allows a finite number of stacky points. \todo{I don't know if this is a good statement to make.  We aren't really viewing $\sx$ as a scheme.  Rather we're are determining information about its canonical rings through knowledge of scheme data.}  A {\bf log spin curve} is a triple $(\sx, \Delta, \halfcan)$ where $\Delta \in \di X$ is an effective divisor and $2\halfcan = K_X + \Delta + \sum_{i=1}^{r}\subhalf{e_i}P_i.$
The central object of study in this paper is the {\bf log spin
canonical ring} of $(\sx, \Delta, \halfcan)$ is
\begin{align*}
	R(\sx, \Delta, \halfcan) = \bigoplus_{d \geq 0} H^0(\sx,\lfloor d \halfcan \rfloor)
\end{align*}

@@

The goal of this paper is to provide a stricter bound on the
generator and relation degrees of log spin canonical rings
for all genera. Note that O'Dorney's bounds are asymptotically
strict for arbitrary $\BQ$-divisors on $\BP^1$, but restricting
to the log spin curve case allows for lower bounds.
The main theorem of this paper is the following improved bound on
the degree of generators and relations of log spin canonical rings:

\todo{State for even weight modular forms attributed to VZB.}
\begin{thm}
\label{thm:main}
Let $(\sx, \Delta, \halfcan)$ be a log spin curve over a perfect field $k$, \todo{Ask David: do we need this perfectness assumption everywhere? No, use seperable stacky points} so
that $\sx$ has signature $\sigma = (0; e_1, \ldots, e_r; \delta)$. Then the
canonical ring

@@

\ssec{Stacky Curves and Log Spin Canonical Rings}
\label{ssec:stacky-background}
A \textbf{stacky curve} $\sx$ over a field $k$ is a smooth proper
geometrically-connected Deligne-Mumford stack of dimension $1$ over
$k$ with a dense open subscheme. A stacky curve is \textbf{tame} if
its stabilizers are coprime to char $k$. Throughout this paper,
stacky curves will be assumed to be tame and $k$ will always be
algebraically closed. The language of stacks will be omitted in
this paper, as it is sufficient to view stacky curves and related
objects $\sx$ as a scheme whose associated divisor group allows for
a finite number of \textbf{stacky points} which may be viewed
heuristically as ``fractional points''. See Voight and
Zureick-Brown for more thorough treatment of these definitions
\cite[Chapter 5]{vzb:stacky}.

@@

\noindent
When the log spin curve is fixed, we usually use $R$ or $R_L$ to
represent $R(\sx, \Delta, \halfcan)$. We also often use $H^0(\sx, D)
:= H^0(\sx, \sco(D))$ and $h^0(\sx, \sco(D)) := \dim H^0(\sx, \sco(D))$.

@@

If $R$ is a graded ring, then $(R)_d$ refers to the $d^{th}$ degree
component of $R$.

@@

\begin{lem}
\label{lem:sat-2}
Let $(\sx, \Delta, \halfcan_\sx)$ and $(\sx', \Delta, \halfcan_{\sx'})$ \todo{The notation i was using for these throughout the paper was $\halfcan$ and $\halfcan'$ without the $\sx$ subscripts. We should make one of them consistent with the other.} be log spin curves with the same coarse space $X = X'$ having signatures $(g; e_1, \ldots, e_r; \delta)$ and $(g, e_1, \ldots, e_{r- 1}, \delta)$, where $e_r=3$.  Suppose $g > 0,$ and, if $g = 1,$ then $\deg 3\halfcan' \geq 2.$ Then, $\sat(Eff(L'))\le 2$. \todo{Why is it automatic that the saturation is 2? Maybe you can explain this in the proof in a sentence.}
Furthermore, let $R_{L'} = k[x_2, x_3
, x_5, \ldots, x_m]/I'$ and let $\halfcan = \halfcan' + \frac{1}{
3}P$, where $P\in X$ is a base point of $\halfcan'$ (which includes the case when $H^0(X, \lfloor \halfcan'\rfloor) = 0$). \todo{Ask David, is it useful to include this parenthetical? yes}
Suppose for $i \in \{2, 3\}$, $\deg x_i = i$ and $\ord_P^{\halfcan'}(x_i)= 0$ (which is satisfied generically in degree 2 and 3, since base-point freeness follows by Riemann-Roch \todo{Probably include this in a remark below the proof}). Choose an ordering on $k[x_2, \ldots, x_m]$ that satisfies
\begin{align*}
	\ord_{x_2}(f) < \ord_{x_2}(h) \implies f \prec h.
\end{align*}

@@

Therefore, general elements $y_3$ and $y_4$ satisfy $-\ord_P^{\halfcan'}(y_i) = 1$ by Riemann-Roch, proving part (a). \todo{Ask David, Understand why/whether we need to cite Lemma 5.4.7. we don't just say it at the beginning}

@@

\iffalse
\begin{rem}
Note that $S_{\sigma,J}(i,d-1)$ as defined in condition (Ad-iii) of Definition ~\ref{defn:admissible} can equivalently be defined as $S_{\sigma,J}(i,d-1)= \{j \in J : j \neq i \text{ and }(e_j' + 2d-2)|e_i' + 2d\}.$ Written this way, $\# S_{\sigma,J}(i,d-1)$ is more apparently similar to $\mu(i,d)$ as defined in \cite[Defintion 8.5.1]{vzb:stacky}.
\end{rem}
\todo{Peter: I'm not sure how useful this remark is.  After seeing the definition of $S_{\sigma, J}(I,d)$ a reader will probably be a little mystified why it is a useful condition.  The inclusion of the relationship to something defined in VZB that also seems ad hoc and that they probably will not have seen may not be the most helpful use of this space.}
\fi

@@

\begin{lem}
\label{lem:genus-2-generation-5}
Let $X$ be a smooth genus 2 curve with canonical divisor $K$ and let $L \in \di X$ be so that $2L \sim K$. 
Then, $R_L$ is generated in degree at most $5$.
\end{lem}
\begin{proof}
By Lemma ~\ref{lem:semicanonical_generation} it suffices to show $R_L$ has no minimal generators in degree 6.
Let $P_1, \ldots, P_6$ be the six hyperelliptic fixed points of $X$. 
Recall that all linear equivalence classes of spin divisors on $X$ have a representative of the form
$$L = P_1 + \sum_{i =2}^{6} \epsilon_i (P_1 - P_i),$$ 
with $\epsilon_i \in \{0,1\},$ and not all $\epsilon_i = 1$. \todo{cite this somewhere, maybe ask david for a place to cite? We can cite http://www.math.harvard.edu/~ctm/home/text/class/harvard/213b/14/html/home/course/course.pdf, for example, from complex analysis course notes! Don't use coarse notes}
Note that $H^0(X,K)$ is spanned by rational functions $u, x$ with $\di u|_{2L} = -2L, \di x|_{2L} = -2L + 2P_1,$ where $u,x$ are though of as rational functions. Therefore, the image of the multiplication map 
\begin{align*}
	Sym^3 H^0(X,2L) \rightarrow H^0(X,6L)
\end{align*}
is 4 dimensional by Riemann--Roch, with image spanned by $u^3,u^2x,ux^2, x^3$, 
which have 
\begin{align*}
	\di u^3|_{2L} &= -2L \\
	\di u^2x|_{2L} &= -2L + 2P_1 \\
	\di ux^2|_{2L} &= -2L+4P_1, \\
	\di x^3|_{2L} &= -2L + 6P_1.
\end{align*}

@@

\begin{align*}
	\dim_k (k[x_{3, 1}, x_{5, 1}, x_{7, 1}, x_{7, 2}] / I)_d &=
	\dim_k (k[x_{3, 1}, x_{5, 1}, x_{7, 1}, x_{7, 2}]	/ in_\prec(I))_d \\
	&= \dim_k (k[x_{3, 1}, x_{5, 1}, x_{7, 1}, x_{7, 2}] / \langle x_{5, 1}^2,
	x_{7, 1}^2 \rangle
\end{align*}
